Insert your title here

### 2018-08-14 sect5-upper-bounds.tex
MI: Simplification of notation of Alg. 2 and a consistent update of Alg. 3. Fixed an error in constraint (1d). Constraints (4d) and (4e) updated to their stronger versions and the explanation in the text also changed. Although it is rather a description of what is written mathematically. Notation N(U) introduced in Sect. 2.

A knowledge of an upper bound $\bar{t}$ affects the number of variables in all studied models. 
Particularly in the decision versions, the iterative approach can be terminated once the solution is not satisfactory for broadcast time limit $\bar{t}-1$.
The algorithms presented in this section iteratively construct broadcast forest $T_i, i=1,2,\dots$.
\subsection{Matching heuristic}

A fast straightforward algorithm determining an upper bound is based on repetitive construction of a matching between already connected nodes and their neighbors.
Initially, the constructed broadcast forest consists of isolated sources. 
In each iteration, the nodes already in the forest are sorted according to the number of neighbors not yet added to the forest.
For each node we then select a neighbor to be appended to the forest.
Let $T_s=(V_s,A_s)$ be the tree rooted at source $s\in S$, and let $T=(V_T, A_T)=\{T_s:s\in S\}$ be the broadcast forest.
Further, nodest of the broadcast forest are numbered by integers $V_T=\{v_1,\dots,v_{|V_T|}\}$.
Alg. \ref{alg:matchheur} formalizes the steps.
%\begin{algorithm}[]
%\KwData{$G=(V,E), S\subseteq V$}
%\textbf{for }$s\in S\textbf{ do } V_s\leftarrow \{s\}, A_s\leftarrow\emptyset$\;
%$t\leftarrow 0$\;
%\While{$V_T\neq V$} {
%	Sort nodes in $V_T$ increasingly by the number of neighbors in $V\setminus V_T$\;
%	$W\leftarrow\emptyset$\;
%	\For{$u\in V_T$} {
%		Find $v\in V\setminus (V_T\cup W)$ such that $(u,v)\in A$\;
%		$W\leftarrow W\cup\{v\}$\;
%	}
%	$A_T\leftarrow A_T\cup \{\{u,w\}: u\in V_T, w\in N(u)\cap W\}$\;
%	$V_T\leftarrow V_T\cup W$\;
%	$t\leftarrow t+1$\;
%}
%\Return t\;
%%\Return $\lceil k/s \rceil$\;
% \caption{A greedy matching heuristic method for determining an upper bound}
%\label{alg:matchheur}
%\end{algorithm}

\begin{algorithm}[]
\KwData{$G=(V,E), S\subseteq V$}
\textbf{for }$s\in S\textbf{ do } V_s\leftarrow \{s\}, A_s\leftarrow\emptyset$\;
$t\leftarrow 0$\;
\While{$V_T\neq V$} {
	Sort nodes in $V_T$ so that $i<j\Leftrightarrow |N(v_i)\cap(V\setminus V_T)|<|N(v_j)\cap (V\setminus V_T)|$\;
	$W\leftarrow\emptyset$\;
	$D\leftarrow\emptyset$\;
	\For{$i=1,\dots,|V_T|$} {
		Find $v_j\in V\setminus (V_T\cup W)$ such that $(v_i,v_j)\in A$\;
		$W\leftarrow W\cup\{v_j\}$\;
		$D\leftarrow D\cup\{\left(v_i,v_j\right)\}$\;
	}
	$V_T\leftarrow V_T\cup W$\;
	$A_T\leftarrow A_T\cup D$\;
	$t\leftarrow t+1$\;
}
\Return t\;
%\Return $\lceil k/s \rceil$\;
 \caption{A greedy matching heuristic method for determining an upper bound}
\label{alg:matchheur}
\end{algorithm}
The selection of a neighbor (line 7) can be done arbitrarily, or according to some strategy such as neighbor with maximum degree is selected first.
Note that for determining an upper bound, it is not necessary to store arcs sets $A_T$ and $D$, because we merely count the number of iterations.
In order to construct an actual solution with broadcast time equal to the upper bound, the arcs sets are needed.
Matchings created in each iteration are not necessarily maximum.

@@

The following inexact method is based on the idea of finding maximum cardinality matching in $G\left[V_{T_i},V\setminus V_{T_i}\right]$, and extending $T_i$ by this matching.
It means that in each iteration $i$, the maximum possible number of nodes are informed.
The maximum cardinality matching can be regarded as finding $|V_{T_i}|$ node-disjoint  binomial trees of order at most one with roots in $V_{T_i}$, maximizing the number of edges.
By generalizing this notion, we iteratively look for $|V_{T_i}|$ node-disjoint (pruned) binomial trees of an arbitrary given order $k$ valued between 1 and $n-|S|$.
Even though this problem is NP-hard for $k\geq 2$ \cite{jansen95}, it is expected that the computational time is sound in most practical instances.
After obtaining the set of binomial trees, first $p$ nodes in each tree are selected and added to the broadcast forest.
The parameter $p\in \{1,\dots,2^k\}$ is also a part of the input.
\begin{algorithm}[]
\KwData{$G=(V,E), S\subseteq V, k\in \{1,\dots,n-|S|\}, p\in \{1,\dots,2^k\}$}
\textbf{for }$s\in S\textbf{ do }T_s=(V_s,A_s), V_s\leftarrow \{s\}, E_s\leftarrow\emptyset$\;
$I=\{1,\dots,2^k\}$\\
$T=\{T_s:s\in S\}$\;
$t\leftarrow 0$\;
\While{$V_T\neq V$} {
	$S\leftarrow V_T$\;
	Find a set of pruned binomial trees $B=\{B_1,\dots,B_{|V_T|}\}$ of order at most $k$ with roots in $V_T$ by solving model \eqref{mod:genmatch}\;
	$V_T\leftarrow V_T\cup \{v:v\in V_B:\beta(v)\leq p\}$\;
	$E_T\leftarrow E_T\cup \{\{u,v\}\in E_B: \beta(u)\leq p,\beta(v)\leq p\}$\;
	$t\leftarrow t+1$\;
}
\Return t\;
%\Return $\lceil k/s \rceil$\;
 \caption{A method for determining an upper bound based on iterative search for pruned binomial trees}
\label{alg:match}
\end{algorithm}

@@

Not only determine the algorithms the upper bound on the minimum broadcast time, but they also construct a solution with this broadcast time.

>>>>

A knowledge of an upper bound $\bar{t}$ affects the number of variables in all studied models. 
Particularly in the decision versions, the iterative approach can be terminated once the solution is found to be infeasible for broadcast time limit $\bar{t}-1$.
The algorithms presented in this section iteratively construct broadcast forest $T=(V_T,A_T)$, where in the last iteration $V_T=V$.
\subsection{Matching heuristic}

A fast straightforward algorithm determining an upper bound is based on repetitive construction of a matching between already connected nodes and their neighbors.
Initially, the constructed broadcast forest consists of isolated sources. 
In each iteration, the nodes already in the forest are sorted according to the number of neighbors not yet added to the forest.
For each node we then select a neighbor to be appended to the forest.
Let $V_T=\{v_1,\dots,v_{|V_T|}\}$.
The steps are formalized in Alg. \ref{alg:matchheur}.
%\begin{algorithm}[]
%\KwData{$G=(V,E), S\subseteq V$}
%\textbf{for }$s\in S\textbf{ do } V_s\leftarrow \{s\}, A_s\leftarrow\emptyset$\;
%$t\leftarrow 0$\;
%\While{$V_T\neq V$} {
%	Sort nodes in $V_T$ increasingly by the number of neighbors in $V\setminus V_T$\;
%	$W\leftarrow\emptyset$\;
%	\For{$u\in V_T$} {
%		Find $v\in V\setminus (V_T\cup W)$ such that $(u,v)\in A$\;
%		$W\leftarrow W\cup\{v\}$\;
%	}
%	$A_T\leftarrow A_T\cup \{\{u,w\}: u\in V_T, w\in N(u)\cap W\}$\;
%	$V_T\leftarrow V_T\cup W$\;
%	$t\leftarrow t+1$\;
%}
%\Return t\;
%%\Return $\lceil k/s \rceil$\;
% \caption{A greedy matching heuristic method for determining an upper bound}
%\label{alg:matchheur}
%\end{algorithm}

\begin{algorithm}[]
\KwData{$G=(V,E), S\subseteq V$}
$V_T\leftarrow S, A_T \leftarrow \emptyset$\;
$\bar{t}\leftarrow 0$\;
\While{$V_T\neq V$} {
	Sort nodes in $V_T$ so that $i<j\Leftrightarrow |N(v_i)\cap(V\setminus V_T)|<|N(v_j)\cap (V\setminus V_T)|$\;
	$W\leftarrow\emptyset$\;
	$D\leftarrow\emptyset$\;
	\For{$i=1,\dots,|V_T|$} {
		Find $v_j\in V\setminus (V_T\cup W)$ such that $(v_i,v_j)\in A$\;
		$W\leftarrow W\cup\{v_j\}$\;
		$D\leftarrow D\cup\{\left(v_i,v_j\right)\}$\;
	}
	$V_T\leftarrow V_T\cup W$\;
	$A_T\leftarrow A_T\cup D$\;
	$\bar{t}\leftarrow \bar{t}+1$\;
}
\Return $\bar{t}$\;
%\Return $\lceil k/s \rceil$\;
 \caption{A greedy matching heuristic method for determining an upper bound}
\label{alg:matchheur}
\end{algorithm}
The selection of a neighbor (line 7) can be done arbitrarily, or according to some strategy such as neighbor with maximum degree is selected first.
Note that for determining an upper bound, it is not necessary to store arcs sets $A_T$ and $D$, because we merely count the number of iterations.
In order to construct an actual solution with broadcast time equal to the upper bound, the arcs sets are needed.
Matchings created in each iteration are not necessarily maximum.

@@

The following inexact method is based on the idea of finding maximum cardinality matching in $G\left[V_T,V\setminus V_T\right]$, and extending $T$ by this matching.
It means that in each iteration, the maximum possible number of nodes are informed.
The maximum cardinality matching can be regarded as finding $|V_T|$ node-disjoint  binomial trees of order at most one with roots in $V_T$, maximizing the number of edges.
By generalizing this notion, we iteratively look for $|V_T|$ node-disjoint (pruned) binomial trees of an arbitrary given order $k$ valued between 1 and $n-|S|$, but typically not more than 4.
Even though this problem is NP-hard for $k\geq 2$ \cite{jansen95}, it is expected that the computational time is sound in most practical instances.
After obtaining the set of binomial trees, first $p$ nodes in each tree are selected and added to the broadcast forest.
The parameter $p\in \{1,\dots,2^k\}$ is also a part of the input.
\begin{algorithm}[]
\KwData{$G=(V,E), S\subseteq V, k\in \{1,\dots,n-|S|\}, p\in \{1,\dots,2^k\}$}
$V_T\leftarrow S, A_T \leftarrow \emptyset$\;
$\bar{t}\leftarrow 0$\;
\While{$V_T\neq V$} {
	$S\leftarrow V_T$\;
	Find a set of pruned binomial trees $B=\{B_1,\dots,B_{|V_T|}\}$ of order at most $k$ with roots in $V_T$ by solving model \eqref{mod:genmatch}\;
	$V_T\leftarrow V_T\cup \{v:v\in V_B:\beta(v)\leq p\}$\;
	$A_T\leftarrow A_T\cup \{(u,v)\in A_B: \beta(u)\leq p,\beta(v)\leq p\}$\;
	$\bar{t}\leftarrow \bar{t}+1$\;
}
\Return $\bar{t}$\;
%\Return $\lceil k/s \rceil$\;
 \caption{A method for determining an upper bound based on iterative search for pruned binomial trees}
\label{alg:match}
\end{algorithm}

@@

Like in Alg. \ref{alg:matchheur}, for calculating an upper bound, it is also not needed to store node set $A_T$ in Alg \ref{alg:match}, but it is essential when the actual broadcast tree is desirable.
